Issues with Dynamical Measurements Across the Disk

### 2019-02-26 paper/ms.tex
Update abstract

Stellar actions, computed for stars from their 6D phase space measurement and
an assumed Galactic potential, are used to label orbits. In principle, actions
have the attractive quality of being invariant and are thus good labels of a
star's orbit. Of course, inaccurate measurements of the 6D phase space
position and uncertainities in the assumed Galactic potential induce an error
in the computed action. We show that, in addition to these complications, a
systematic bias in the 6D phase space position induces a phase-dependence in
the actions, which we interpret as an error. This error is non-Gaussian and
bimodal, with neither of the modes peaking on the true action value. A
midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to induce a
$100\,\%$ error in the vertical action $J_z$ for thin and thick disk orbits,
respectively. Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in cosmological zoom-in simulations from
the FIRE collaboration. Thus, current state-of-the-art action calculations ---
which assume the global and local midplanes are the same --- include a
systematic offset in $z$. Variation in the local standard of rest will induce
similar issues. The variation of the midplane must be taken into account when
performing dynamical modelling across large regions of the disk accessible
with the {\em Gaia} mission.

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
offset coordinate systems depend on the time (i.e. orbital phase) at which the
star's phase-space position is observed. (To our extremely long-lived observer
this would appear as time-dependence in the actions, which appear to oscillate
around their value using the correct coordinate system even though in this
example the Galaxy is described perfectly by a static, axisymmetric
potential.) The relative size of the phase variation in each action depends on
the direction of the systematic offset as well as the true values of the
actions (i.e. the type of orbit). In reality we will have one measurement of
the phase-space position to work with, in which case the determination of the
orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
in that coordinate (see Figure \ref{fig:cartoon}). Additionally, in searching
for migrated members of known open clusters, we would compute actions for
stars with a wide range of orbital phases and locations in the Galaxy. In the
following we therefore quote percentile ranges for the possible values
computed for each action as a proxy for the effect of these systematic errors
in the coordinate system.

@@

To illustrate this concept, we first compute the absolute range of midplane
values for a given $\Delta \phi$ segment of the midplane. With the observer's
azimuth set to zero, we consider a region extending from $-\Delta \phi/2$ to
$\Delta\phi/2$. To translate this angular scale into a physical scale, we also
compute the chord length $\lambda$ for the chord extending from the observer
to $\pm \Delta\phi/2$ (see Figure \ref{fig:fig_to_explain}). We repeat this
for each of $50$ bins in azimuth. Each bin is plotted in
Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid
blue and the $1\,\sigma$ dispersion as the dashed blue lines.

>>>>

Stellar actions, computed for stars from their 6D phase space measurement and
an assumed Galactic potential, are used to label and distinguish orbits. In
principle, actions have the attractive quality of being invariant and are thus
good labels of a star's orbit. Of course, inaccurate measurements of the 6D
phase space position and uncertainities in the assumed Galactic potential
induce an error in the computed action. We show that, in addition to these
complications, a systematic bias in the 6D phase space position induces a
phase-dependence in the actions, which we interpret as an error. This error is
non-Gaussian and bimodal, with neither of the modes peaking on the true action
value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to
induce a $100\,\%$ error in the vertical action $J_z$ for thin and thick disk
orbits, respectively. Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
simulations from the FIRE collaboration. Thus, current state-of-the-art action
calculations --- which assume the global and local midplanes are the same ---
include a systematic offset in $z$. Variation in the local standard of rest
will induce similar issues. The variation of the midplane must be taken into
account when performing dynamical modelling across large regions of the disk
accessible with the {\em Gaia} mission.

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
offset coordinate systems depend on the time (i.e. orbital phase) at which the
star's phase-space position is observed. (To our extremely long-lived observer
this would appear as time-dependence in the actions, which appear to oscillate
around their value using the correct coordinate system even though in this
example the Galaxy is described perfectly by a static, axisymmetric
potential.) The relative size of the phase variation in each action depends on
the direction of the systematic offset as well as the true values of the
actions (i.e. the type of orbit). In reality we will have one measurement of
the phase-space position to work with, in which case the determination of the
orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
in that coordinate (see Figure~\ref{fig:cartoon}). Additionally, in searching
for migrated members of known open clusters, we would compute actions for
stars with a wide range of orbital phases and locations in the Galaxy. In the
following we therefore quote percentile ranges for the possible values
computed for each action as a proxy for the effect of these systematic errors
in the coordinate system.

@@

To illustrate this concept, we first compute the absolute range of midplane
values for a given $\Delta \phi$ segment of the midplane. With the observer's
azimuth set to zero, we consider a region extending from $-\Delta \phi/2$ to
$\Delta\phi/2$. To translate this angular scale into a physical scale, we also
compute the chord length $\lambda$ for the chord extending from the observer
to $\pm \Delta\phi/2$ (see Figure~\ref{fig:fig_to_explain}). We repeat this
for each of $50$ bins in azimuth. Each bin is plotted in
Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid
blue and the $1\,\sigma$ dispersion as the dashed blue lines.
